Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-02-13 pkg/disttree/inst/draft/draft.Rnw
disttree and distforest without family list stored in inner nodes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2014 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 38e2320ef5303575cdb1185228fee02dbd65ece3

# dist_list_cens_normal
{

  dist_list_cens_normal <- list()

  parnames <- c("mu", "sigma")
  etanames <- c("mu", "log(sigma)")

  ddist <-  function(y, eta, log = TRUE, weights = NULL, sum = FALSE, left = 0, right = Inf) {     
    par <- c(eta[1], exp(eta[2]))
    val <- crch::dcnorm(x = y, mean = par[1], sd = par[2], left = left, right = right, log = log)
    if(sum) {
      if(is.null(weights)) weights <- if(is.matrix(y)) rep.int(1, dim(y)[1]) else rep.int(1, length(y))
      val <- sum(weights * val, na.rm = TRUE)
    }
    return(val)
  }

  sdist <- function(y, eta, weights = NULL, sum = FALSE, left = 0, right = Inf) {   
    par <- c(eta[1], exp(eta[2]))
    # y[y==0] <- 1e-323

    score_m <- crch:::scnorm(x = y, mean = par[1], sd = par[2], which = "mu", left = left, right = right)
    score_s <- crch:::scnorm(x = y, mean = par[1], sd = par[2], which = "sigma", left = left, right = right) * exp(eta[2]) # inner derivation exp(eta[2])
    score <- cbind(score_m, score_s)
    score <- as.matrix(score)
    colnames(score) <- etanames
    if(sum) {
      if(is.null(weights)) weights <- rep.int(1, length(y)[1])
      # if score == Inf replace score with 1.7e308 because Inf*0 would lead to NaN (0 in weights)
      score[score==Inf] = 1.7e308
      score <- colSums(weights * score, na.rm = TRUE)
      #if(any(is.nan(score))) print(c(eta, "y", y))
    }
    return(score)
  }

  hdist <- function(y, eta, weights = NULL, left = 0, right = Inf) {    
    ny <- length(y)
    if(is.null(weights)) weights <- rep.int(1, ny)

    par <- c(eta[1], exp(eta[2]))                           
    # y[y==0] <- 1e-323

    d2mu <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "mu", left = left, right = right)
    d2sigma <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "sigma", left = left, right = right)
    dmudsigma <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "mu.sigma", left = left, right = right) # FIX: order?
    dsigmadmu <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "sigma.mu", left = left, right = right) # FIX: order?
    dsigma <- crch:::scnorm(x = y, mean = par[1], sd = par[2], which = "sigma", left = left, right = right)

    d2ld.etamu2 <- sum(weights * d2mu, na.rm = TRUE)
    d2ld.etamu.d.etasigma <- sum(weights * dmudsigma * par[2], na.rm = TRUE)
    d2ld.etasigma.d.etamu <- sum(weights * dsigmadmu * par[2], na.rm = TRUE)
    d2ld.etasigma2 <- sum(weights * (d2sigma * exp(2*eta[2]) + dsigma * par[2]), na.rm = TRUE)

    hess <- matrix(c(d2ld.etamu2, d2ld.etamu.d.etasigma, d2ld.etasigma.d.etamu, d2ld.etasigma2), nrow = 2)
    colnames(hess) <- rownames(hess) <-  etanames

    return(hess)
  }

  ## additional functions pdist, qdist, rdist
  pdist <- function(q, eta, lower.tail = TRUE, log.p = FALSE) crch:::pcnorm(q, mean = eta[1], sd = eta[2], 
                                                                            lower.tail = lower.tail, log.p = log.p, 
                                                                            left = left, right = right)
  qdist <- function(p, eta, lower.tail = TRUE, log.p = FALSE) crch:::qcnorm(p, mean = eta[1], sd = eta[2], 
                                                                            lower.tail = lower.tail, log.p = log.p, 
                                                                            left = left, right = right)
  rdist <- function(n, eta) crch:::rcnorm(n, mean = eta[1], sd = eta[2], left = left, right = right)

  link <- c("identity", "log")

  linkfun <- function(par) {
    eta <- c(par[1], log(par[2]))
    names(eta) <- etanames
    return(eta)
  }

  linkinv <- function(eta) {
    par <- c(eta[1], exp(eta[2]))
    names(par) <- parnames
    return(par)
  }

  linkinvdr <- function(eta) {
    dpardeta <- c(1, exp(eta[2]))
    names(dpardeta) <- parnames
    return(dpardeta)
  }

  startfun <- function(y, weights = NULL){
    yc <- pmax(0,y)  # optional ?
    if(is.null(weights)) {
      mu <- mean(yc)
      sigma <- sqrt(1/length(yc) * sum((yc - mu)^2))
    } else {
      mu <- weighted.mean(yc, weights)
      sigma <- sqrt(1/sum(weights) * sum(weights * (yc - mu)^2))
    }
    starteta <- c(mu, log(sigma))
    names(starteta) <- etanames
    return(starteta)
  }

  mle <- FALSE

  dist_list_cens_normal <- list(family.name = "censored Normal Distribution",
                                ddist = ddist, 
                                sdist = sdist, 
                                hdist = hdist,
                                pdist = pdist,
                                qdist = qdist,
                                rdist = rdist,
                                link = link, 
                                linkfun = linkfun, 
                                linkinv = linkinv, 
                                linkinvdr = linkinvdr,
                                startfun = startfun,
                                mle = mle,
                                gamlssobj = FALSE,
                                censored = TRUE
  )
}

@@

%% late: introduce MOS and EMOS? more details about precipitation forecasting
%In ensemble model output statistics (EMOS, \cite{gneiting2005calibrated}) this 
%idea is taken one step further. Usually acting on the assumption of a normally distributed
%response variable this model allows for both parameters, the mean and the variance, to depend 
%linearly on covariates provided by NWPs. However, the restriction to model only linear effects 
%has risen the interest in applying more flexible models such as offered by the GAMLSS framework.

>>>>

@@

%% late: introduce MOS and EMOS? more details about precipitation forecasting
%In ensemble model output statistics (EMOS, \cite{gneiting2005calibrated}) this idea is taken one step further. Usually acting on the assumption of a normally distributed response variable this model allows for both parameters, the mean and the variance, to depend linearly on covariates provided by NWPs. However, the restriction to model only linear effects has risen the interest in applying more flexible models such as offered by the GAMLSS framework.
